---
title: "Research Project"
output:
  html_document: default
  html_notebook: default
---
 

#### Load Packages
```{r warning=FALSE, message=FALSE}
library(quantmod)
library(tidyverse)
library(XLConnect)
library(tidyquant)
library(tseries)
library(reshape2)
library(readxl)
```

####  Function: Calculate monthly return
```{r warning=FALSE, message=FALSE}
get_return <- function(symbol, start, end, compression = 'm'){
        price <- get.hist.quote(instrument = symbol, start = start, end = end, compression = compression, quote = "Close", quiet = TRUE)
        Date <- index(price)

        price <- (price %>% as.data.frame()
                  %>% mutate(Date = Date)
                  %>% select(Date, everything())
        )

        returns <- price %>% mutate(returns = Close/lag(Close) - 1) %>% select(-Close) %>% filter(complete.cases(returns))

        returns <- returns %>% as.tibble()
        return(returns)
}

```

####  Data range and query stock
```{r warning=FALSE, message=FALSE}
start <- '2012-06-01'

end <- '2018-06-02'

query_stock <- 'APA'
```

###  I. Carhart four-factor model

#### 1) Query stock returns
```{r warning=FALSE, message=FALSE}
returns <- get_return(query_stock, start, end)

returns$Date <- as.yearmon(returns$Date)
```

#### 2) Data: Fama French 3 factors
```{r warning=FALSE, message=FALSE}
ff_factors <- read.csv("F-F_Research_Data_Factors.CSV", header = TRUE, skip = 3, nrows = 1102)

colnames(ff_factors)[1] <- 'Date'

# Format Date into yearmon

ff_factors$Date <- as.yearmon(as.Date(paste(ff_factors$Date, '01', sep = ''), format='%Y%m%d'))

ff_factors[,-1] <- ff_factors[,-1]/100
```

#### 3) Data: Fama French Momentum
```{r warning=FALSE, message=FALSE}

ff_mom <- read.csv("F-F_Momentum_Factor.CSV", header = TRUE, skip = 13, nrows = 1095)

colnames(ff_mom)[1] <- 'Date'

# Format Date into yearmon

ff_mom$Date <- as.yearmon(as.Date(paste(ff_mom$Date, '01', sep = ''), format='%Y%m%d'))

ff_mom[,-1] <- ff_mom[,-1]/100
```

#### 4) Run Regression
```{r warning=FALSE, message=FALSE}
# Merge stock returns, ff_mom, ff_factors by Date

df <- Reduce(function(x,y) merge(x, y, by = 'Date'), list(ff_factors, ff_mom, returns))

# Calculate excess stock return

df <- df %>% mutate(Ex.return = returns - RF) 

# Run regression

df <- df %>% select(-c(RF, returns))

summary(lm(Ex.return ~ ., data = df[,-1]))
```

#### >> According to the Carhart four-factor model, momentum and market factor are significant.

### II. PCA Analysis

#### 1) Filter out stocks in the same sub-sector with query_stock. (Data: GICS Subsector)
        
```{r warning=FALSE, message=FALSE}
sub_industry <- read_xlsx("SP500 Sectors.xlsx", sheet = 1) 

```

#### 2) Sub-industry for query stock

```{r warning=FALSE, message=FALSE}
query_industry <- sub_industry %>% filter(`Ticker symbol` == query_stock) %>% select(`GICS Sub Industry`) %>% unlist()

query_industry %>% as.character()
```


#### 3) Filter out peers

```{r warning=FALSE, message=FALSE}
peers <- sub_industry %>% filter(`GICS Sub Industry` == query_industry) %>% select(`Ticker symbol`)

colnames(peers) <- 'symbol'

peers %>% unlist() %>% as.vector()
```

#### 4) Calculate returns for stocks in the selected sub-sector
```{r warning=FALSE, message=FALSE}
peers_return <- peers %>% mutate(returns = map(symbol, function(.x) get_return(.x, start, end))) 

peers_return <- peers_return %>% unnest()

peers_return <- dcast(peers_return, Date ~ symbol)

# Exclude stocks with shorter history than the data range selected 
peers_return <- peers_return[, colSums(is.na(peers_return)) == 0]
```

#### 5) PCA analysis
```{r warning=FALSE, message=FALSE}
pca <- prcomp(peers_return[,-1])
```

#### 6) Select retained PCs based on Scaled Average Eigenvalues (Keep eigenvalues larger than 0.7*mean(eigenvalues))
```{r warning=FALSE, message=FALSE}
keep_scale <- mean(pca$sdev^2)*0.7
n <- sum(pca$sdev^2 > keep_scale)
n
```

#### 7) Regress APA stock return on PCs
```{r warning=FALSE, message=FALSE}
pcs <- pca$x[,c(1:n)]

pc_tbl <- peers_return %>% select(Date) %>% cbind(pcs)

pc_tbl$Date <- as.yearmon(pc_tbl$Date)

tbl <- merge(returns, pc_tbl)

summary(lm(returns ~., data = tbl[,-1]))
```

#### >> PC1 and PC3 are significant, R-squared is 0.6884.

#### 8) Variance explained by PCA
```{r warning=FALSE, message=FALSE}
summary(pca)$importance[,1:3]
```

#### >> First three PCs have explained 78.85% of variation

#### 9) PC1, PC2, PC3 loadings
```{r warning=FALSE, message=FALSE}
pca$rotation[,1:n]
```

####>> All stocks have positive weights in PC1, check its correlation with Oil Price
####10) Data: WTI
```{r warning=FALSE, message=FALSE}
WTI <- read.csv('WTI.csv', header = TRUE)

WTI$Date <- as.yearmon(as.Date(WTI$Date, format='%m/%d/%Y'))

WTI <- WTI %>% mutate(WTI.return = WTI/lag(WTI)-1) %>% select(-WTI)
```

#### 11) Correlation tabel
```{r warning=FALSE, message=FALSE}
df_new <- merge(tbl, WTI)

cor(df_new[,-1])
```

#### >> The correlation betwwen PC1 and WTI is 0.44, because the sub-industry of APA is Oil & Gas Exploration & Production, it may have higher correlation with Oil & Gas index, commodity index or PMI. (But those data cannot be downloaded from website, it is not feasible to verify this guess)

### III. Further thoughts

#### >> Because this poject is to study the return of one single stock, a model based on valuation variables would explain more firm-specific returns, factors including Size, P/B, ROE, Dividends per share, Debt/Price, etc. But those historical valuation data cannot be obtained online, so it is not feasible to do more research on this idea.
