---
title: "Research Project"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---

```{r include=FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

#### Load Packages
```{r warning=FALSE, message=FALSE}

library(quantmod)
library(tidyverse)
library(XLConnect)
library(tseries)
library(car)

```

#### Function: Obtain Prices

```{r}

get_prc <- function(symbol, start, end, compression = 'm'){
        
        price <- get.hist.quote(instrument = symbol, start = start, end = end, compression = compression, quote = "Close", quiet = TRUE)
        
        Date <- index(price)

        price <- (price %>% as.data.frame()
                  %>% mutate(Date = Date)
                  %>% select(Date, everything())
        )
        
        return(price)
}

```

####  Function: Calculate Monthly Returns

```{r warning=FALSE, message=FALSE}

get_return <- function(symbol, start, end, compression = 'm'){
        
        price <- get.hist.quote(instrument = symbol, start = start, end = end, compression = compression, quote = "Close", quiet = TRUE)
        
        Date <- index(price)

        price <- (price %>% as.data.frame()
                  %>% mutate(Date = Date)
                  %>% select(Date, everything())
        )

        returns <- price %>% mutate(returns = Close/lag(Close) - 1) %>% select(-Close) %>% filter(complete.cases(returns))

        returns <- returns %>% as.tibble()
        
        return(returns)
}

```


#### Function: Variance Inflation Factor Test

```{r warning=FALSE, message=FALSE}

VIF_test <- function(df, y){
        
        s <- 11
        
        # If VIF > 10, exclude factor
        
        while(sum(s > 10) != 0){ 
                
                model <- lm(df[,y] ~ ., data = df[,names(df) != y])
                
                s <- vif(model)
                
                df <- df %>% select(-one_of(names(s[s >10])))
                
        }
        
        # Remove factor with VIF > 4 after factors of high collinearity (VIF > 10) has been removed
        
        while(sum(s > 4) != 0){
                
                model <- lm(df[,y] ~ ., data = df[,names(df) != y])
                
                s <- vif(model)
                
                df <- df %>% select(-one_of(names(s[s >4])))
                
        }
        
        return(df)
        
}

```

####  Data range and query stock
```{r warning=FALSE, message=FALSE}

start <- '1998-12-31'

end <- '2018-06-02'

query_stock <- 'MSFT'

```

###  I. Carhart four-factor model

#### 1) Obtain stock returns
```{r warning=FALSE, message=FALSE}

returns <- get_return(query_stock, start, end)

returns$Date <- as.yearmon(returns$Date)

```

#### 2) Data: Fama French 3 factors
```{r warning=FALSE, message=FALSE}
ff_factors <- read.csv("F-F_Research_Data_Factors.CSV", header = TRUE, skip = 3, nrows = 1102)

colnames(ff_factors)[1] <- 'Date'

# Format Date into yearmon

ff_factors$Date <- as.yearmon(as.Date(paste(ff_factors$Date, '01', sep = ''), format='%Y%m%d'))

ff_factors[,-1] <- ff_factors[,-1]/100
```

#### 3) Data: Fama French Momentum
```{r warning=FALSE, message=FALSE}

ff_mom <- read.csv("F-F_Momentum_Factor.CSV", header = TRUE, skip = 13, nrows = 1095)

colnames(ff_mom)[1] <- 'Date'

# Format Date into yearmon

ff_mom$Date <- as.yearmon(as.Date(paste(ff_mom$Date, '01', sep = ''), format='%Y%m%d'))

ff_mom[,-1] <- ff_mom[,-1]/100
```

#### 4) Run Rolling Regression
```{r warning=FALSE, message=FALSE}
# Merge stock returns, ff_mom, ff_factors by Date

df <- Reduce(function(x,y) merge(x, y, by = 'Date'), list(ff_factors, ff_mom, returns))

# Calculate excess stock return

df <- df %>% mutate(Ex.return = returns - RF) 

# Run rolling regression with window length: 120 months

df <- df %>% select(-c(RF, returns))

ff_list <- list()

from <- 120

for (i in from:nrow(df)){
        
        # Data within the window
        
        tem_data <- df[(i-from+1):i, -1]
        
        # Regression summary
        
        tem_s <- summary(lm(Ex.return ~., data = tem_data))
        
        # Filter out factors that are significant at 95% confidence level
        
        tem_ff <- rownames(tem_s$coefficients)
        
        ff_list[[i]] <- tem_ff[tem_s$coefficients[,4] < 0.05]
}

# Count significant factors occurence in the recent 60 months

table(do.call(rbind, tail(ff_list, 60)))

```

#### >> According to the Carhart four-factor model, market return and SMB are significant for recent 60 months. Those two factors will be combined with valuation model.

### II. Valuation Model

#### 1) Load Quarterly Fundamentals 

```{r warning=FALSE, message=FALSE}

# Fundamentals data frequenct: Quarterly

fund_qtr <- read.csv("Fundamentals.csv", header = T)

```

#### 2) Fundamental Factors

##### a) Valuation Factors: Price/Sales, Price/Earnings, Dividend Yield

```{r}

Mod2_factors <- fund_qtr %>% mutate(Date = datadate, 
                                    Price_Sales = Price/Sales, 
                                    P_E = Price/NI,
                                    Div_yield = Div/Price)

# Format date

Mod2_factors$Date <- as.Date(as.character(Mod2_factors$Date), 
                             format = '%Y%m%d')

```

##### b) Growth Factors: 3-month Sales Growth, 1-year Sales Growth, 3-month EPS Growth, 1-year EPS Growth

```{r}

Mod2_factors <- Mod2_factors %>% mutate(Sales_g3M = Sales/lag(Sales,1)-1,
                                        Sales_g1Y = Sales/lag(Sales,3)-1,
                                        EPS_g3M = EPS/lag(EPS,1)-1,
                                        EPS_g1Y = EPS/lag(EPS,3)-1)

```

##### c) Quality Factors: Current Ratio, Asset/Equity, Working Capital/Asset

```{r}

Mod2_factors <- Mod2_factors %>% mutate(Current_ratio = CA/CL,
                                        Asset_Equity = TA/TE,
                                        WorkingCap_Asset = (CA-CL)/TA)

```

##### d) Efficiency Factors: Capex/Asset, Capex/Sales, Operating CF/Asset

```{r}

Mod2_factors <- Mod2_factors %>% mutate(Capex_Asset = CapEx/TA,
                                        Capex_Sales = CapEx/Sales,
                                        OptingCF_Asset = OptingCF/TA)

```


##### e) Profitability Factors: ROE, ROA, ROIC
```{r}

Mod2_factors <- Mod2_factors %>% mutate(NI_4Q = rollsum(NI, 4, fill = NA, align = 'right'),
                                        EQ_2Q = rollmean(TE, 2, fill = NA, align = 'right'),
                                        TA_2Q = rollmean(TA, 2, fill = NA, align = 'right'),
                                        IC_2Q = rollmean(IC, 2, fill = NA, align = 'right'),
                                        ROE = NI_4Q/EQ_2Q,
                                        ROA = NI_4Q/TA_2Q,
                                        ROIC = NI_4Q/IC_2Q)

Mod2_factors <- Mod2_factors %>% select(-one_of(c('NI_4Q', 'EQ_2Q', 'TA_2Q', 'IC_2Q')))

```


##### f) Size Factors: Book/Market Value 

```{r}

Mod2_factors <- Mod2_factors %>% mutate(B_M = TE/Mkt.Value)
                                        
```

##### e) Valuation factor table
```{r warning=FALSE, message=FALSE}

tem <- colnames(fund_qtr)

Mod2_factors <- Mod2_factors %>% select(-one_of(tem))

```


#### 3) Price Factors: 6-month Momentum, 1-month Mean Reversion

```{r warning=FALSE, message=FALSE}

# Obtain Prices
# Data Frequency: Monthly 

query_price <- get_prc(query_stock, Mod2_factors$Date[1], Mod2_factors$Date[nrow(Mod2_factors)])

query_price$Date <- query_price$Date-1

# Calculate price factors

prc_factors <- (query_price %>% mutate(Return = Close/lag(Close)-1,
                                       Mom_6M = Close/lag(Close,2)-1,
                                       Avg_12M = rollmean(Return,4, fill = NA, align = 'right'),
                                       Std_12M = rollapply(Return, 4, sd, fill = NA, align = 'right'),
                                       MeanRev_1M = (Return-Avg_12M)/Std_12M)
                            %>% select(Date, Return, Mom_6M, MeanRev_1M))


```


#### 4) Fama-French factors: Market, SMB
```{r warning=FALSE, message=FALSE}

mkt <- ff_factors %>% select(Date, Mkt.RF, SMB, RF)

mkt$Date <- as.Date(mkt$Date, frac = 1)

```


#### 5) Market Risk Factor
```{r warning=FALSE, message=FALSE}

# Market beta from 60-month rolling regression 
# Data Frequency: Monthly

# Obtain market excess returns and risk-free returns

tem1 <- ff_factors %>% select(Date, Mkt.RF, RF)

tem1$Date <- as.Date(tem1$Date, frac = 1)

# Obtain monthly stock returns

tem2 <- get_return(query_stock, tem1$Date[1], tem1$Date[nrow(tem1)])

tem2$Date <- tem2$Date-1

# Calculate stock excess returns

Mod2_df <- merge(tem1, tem2) %>% mutate(Ex.Return = returns - RF) 

# Run rolling regression and obtain betas

betas <- rollapply(data = Mod2_df[,c('Ex.Return','Mkt.RF')], 
                   width = 60, 
                   FUN = function(x) {reg = lm(Ex.Return ~ Mkt.RF, data = as.data.frame(x));
                                      return(as.numeric(reg$coef[2]))},
                   by.column = FALSE, 
                   align = 'right', 
                   fill = NA)     
                          

```

#### 6) Lag Factors for Crtain Period and transform into quarterly data

##### a) Lag valuation factors for two quarter

```{r warning=FALSE, message=FALSE}

Mod2_factors$Date <- lead(Mod2_factors$Date, 2)

```

##### b) Lag price factors for one month

```{r warning=FALSE, message=FALSE}

prc_factors$Date <- lead(prc_factors$Date, 1)

prc_factors <- prc_factors %>% filter(Date %in% Mod2_factors$Date)

```

##### c) Lag market risk factor for one month
```{r}

Mod2_df <- (Mod2_df %>% mutate(beta = lag(betas))
                    %>% filter(Date %in% Mod2_factors$Date))

Mod2_df <- Mod2_df %>% select(Date, beta)

```

##### d) Lag Fama-French factors for one month

```{r}

mkt$Date <- lead(mkt$Date, 1)

mkt <- mkt %>% filter(Date %in% Mod2_factors$Date)

```


#### 5) Aggregate Factor Table

```{r warning=FALSE, message=FALSE}

# Combine valuation factors, price factors, market risk factor and Fama-French factors

factors_all <- Reduce(function(x,y) merge(x, y, by = 'Date'),
                      list(Mod2_factors, prc_factors, Mod2_df, mkt))

```

#### 6) Cope with Collinearity in Valuation Factors with Straightforward Method

```{r}

# Give equal weights to highly correlated factors to create indexes

factors_all <- factors_all %>% mutate(quality = (Current_ratio+Asset_Equity+WorkingCap_Asset)/3,
                                      efficiency = (Capex_Asset+Capex_Sales)/2,
                                      growth1 = (Sales_g3M+Sales_g1Y)/2,
                                      growth2 = (EPS_g3M+EPS_g1Y)/2,
                                      Profitability = (ROA+ROE+ROIC)/3)

# Keep indexes and exclude those factors

exclude_fac <- c('Current_ratio', 'Asset_Equity', 'WorkingCap_Asset', 'Capex_Asset', 'Capex_Sales', 'Sales_g3M', 'Sales_g1Y', 'ROA', 'ROE', 'ROIC', 'EPS_g3M', 'EPS_g1Y')

factors_all <- factors_all %>% select(-one_of(exclude_fac))

```


### III. Factor Analysis of Contemporaneous Returns

```{r warning=FALSE, message=FALSE}

# Calculate excess return of query stock

factors_all <- factors_all %>% mutate(Ex.return = Return-RF) %>% select(-Return, -RF)

output <- list()

# Data window: 60 quarters

from <- 60

# Filter factors with VIF test to avoid collinearity

for (i in from:nrow(factors_all)){
        
        # Data within the window
        
        tem_data <- factors_all[(i-from+1):i, ]
        
        # Select factors with VIF test
        
        tem_factors <- VIF_test(tem_data[,-1], 'Ex.return')
        
        # Output data into list
        
        output[[i]] <- summary(lm(Ex.return ~., data = tem_factors))
        
        # Name list elements by the date of contemporaneous Returns
        
        names(output)[i] <- toString(factors_all$Date[i])
}

# Remove empty lists

output[sapply(output, is.null)] <- NULL

# Print results

output
```

